AN EFFICIENT LOCALIZED MODEL ORDER REDUCTION FRAMEWORK FOR THE SHAPE OPTIMIZATION OF ADDITIVELY MANUFACTURED LATTICE STRUCTURES\break ECCOMAS CONGRESS 2024

### 2024-08-30 paper/paper.tex
add abstract

% TODO abstract
\abstract{\comment[id=pd]{TODO: add abstract}.}

@@

\added[id=pd]{TODO: add note on notation where appropriate.}\\
\added[id=pd]{TODO: shorten introduction.}

@@

% TODO adjust abovecaptionskip if necessary
% it seems the default is 6pt (see class file)
% \setlength{\abovecaptionskip}{15pt plus 3pt minus 2pt}

@@

For the validation of the ROM, the relative error with respect to the FOM is computed over a parameter set of size $200$.
The error in the displacements is measured in the energy norm (over the domain $\varOmega^{\p}_{\gl}$) and defined as
\begin{equation}
    e_u(\m) := \frac{\norm{\bm{u}_{\fom} - \bm{u}_{\rom}}}{\norm{\bm{u}_{\fom}}}\,.
    \label{eq:error_u}
\end{equation}
Apart from the displacement field as the primary unknown, we are interested in the first principal component of the \textsc{Cauchy} stress as this will be used to define a stress constraint for the optimization problem.
The error in the stress is defined as
\begin{equation}
    e_{\sigma}(\m) = \frac{\vert\vb{s}_{\fom} - \vb{s}_{\rom}\vert}{\vert \mathrm{max}(\vb{s}_{\fom})\vert}\,.
    \label{eq:error_sigma}
\end{equation}
Here, $\vb{s}\in\mathbb{R}^{N_g}$ denotes the vector holding the first principal stress values at the $N_g$ integration points of the mesh.
For ease of notation, the parameter-dependence is omitted for $\bm{u}$ and $\vb{s}$.
The minimum, average and maximum of $e_u$ and $e_{\sigma}$ over the validation set are shown in~\cref{fig:relerr}.
With regard to the error in the displacement, the average value is much smaller than the maximum over the validation set.
In both cases, for local basis size ca. $> 75$ the error decreases only very slowly with increasing number of basis functions.
This is due to the rather moderate size of the training set, as \eg{} $N_{\train}=200$ is used for a local parameter space $\mathcal{P}^k=[R_{-}, R_{+}]^4$ for $k=3, \ldots, 9$.
However, the error in the stress is less sensitive to variations in the parameter $\m$ and a pointwise maximum relative error below $1\%$ is acceptable for the current application.
% \comment[id=pd]{Actually, this is probably only because the ROM stress is computed on the full mesh but with reduced $u$ as input. Reduced stress computation?}

>>>>

% TODO abstract
\abstract{A common engineering practice is to optimize the geometry of a structure by an iterative process, in which an objective function is minimized by systematically choosing the value of design variables and computing the value of the objective function many times.
However, regarding multiscale problems, this direct numerical approach is not feasible for a number of reasons.
The model has to take into account both the structural and unit cell (UC) scale, because the objective function to be minimized is a global quantity, while the parameters are related to the shape or material parameters of the UC.
In the absence of a clear separation of scales, the direct resolution of both scales in the numerical model leads to a significant increase in computational cost, which makes it impossible to repeatedly evaluate the model during the optimization process.

The main goal of this contribution is to overcome the aforementioned limitations and develop an efficient computational framework for the optimal design of lattice structures.
To this end, parametric model order reduction is combined with domain decomposition methods.
% The change in geometry of each UC within the lattice is modelled by introducing a transformation that maps the reference to the physical domain.
In offline simulations, suitable localized approximation spaces, that account for the change in geometry of the UC and its neighbours, are constructed using the concept of oversampling and random sampling.
The global approximation space is constructed from the local spaces via the partition of unity method.
The reduced order model is validated on the example of a stress-constrained shape optimization problem.
}

% TODO adjust abovecaptionskip if necessary
% it seems the default is 6pt (see class file)
% \setlength{\abovecaptionskip}{3pt plus 3pt minus 2pt}

@@

\comment[id=pd]{TODO: add note on notation where appropriate.}

\comment[id=pd]{TODO: shorten introduction.}

@@

@@

For the validation of the ROM, the relative error with respect to the FOM is computed over a parameter set of size $200$.
The error in the displacements is measured in the energy norm (over the domain $\varOmega^{\p}_{\gl}$) and defined as
\begin{equation}
    e_u(\m) := \frac{\norm{\bm{u}_{\fom} - \bm{u}_{\rom}}}{\norm{\bm{u}_{\fom}}}\,.
    \label{eq:error_u}
\end{equation}
Apart from the displacement field as the primary unknown, we are interested in the first principal component of the \textsc{Cauchy} stress as this will be used to define a stress constraint for the optimization problem.
The error in the stress is defined as
\begin{equation}
    e_{\sigma}(\m) = \mathrm{max} \frac{\vert\vb{s}_{\fom} - \vb{s}_{\rom}\vert}{\norm{\vb{s}_{\fom}}_2}\,.
    \label{eq:error_sigma}
\end{equation}
Here, $\norm{\cdot}_2$ denotes the Euclidean norm and $\vb{s}\in\mathbb{R}^{N_g}$ denotes the vector holding the first principal stress values at the $N_g$ integration points of the mesh.
For ease of notation, the parameter-dependence is omitted for $\bm{u}$ and $\vb{s}$.
The minimum, average and maximum of $e_u$ and $e_{\sigma}$ over the validation set are shown in~\cref{fig:relerr}.
With regard to the error in the displacement, the average value is much smaller than the maximum over the validation set.
In both cases, for local basis size ca. $> 75$ the error decreases only very slowly with increasing number of basis functions.
This is due to the rather moderate size of the training set, as \eg{} $N_{\train}=200$ is used for a local parameter space $\mathcal{P}^k=[R_{-}, R_{+}]^4$ for $k=3, \ldots, 9$.
However, the error in the stress is less sensitive to variations in the parameter $\m$ and a pointwise maximum relative error below $1\%$ is acceptable for the current application.
% \comment[id=pd]{Actually, this is probably only because the ROM stress is computed on the full mesh but with reduced $u$ as input. Reduced stress computation?}
